Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for

### 2020-09-01 Scripts/Initial_Analysis.Rmd
Fixed Gabs coordinate system and selected LM

#### Project's Library
packages <- c(
  "dbemImport",
  "tidyverse",
  "here", # for dbem_import `here()`
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  "geosphere", # estimate distances between points `distm()`
  "ggrepel",
  "zoo", # for average mean
  "parallel", # for mclapply,
  "sf", # for mapping
  "st", # for mapping
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  "zeallot", # for Juanito's map
  "gmt", # for estimating distances between points
  "viridis",
  "cowplot",
  "rfishbase",
  "stargazer", # For nice lm tables
  "tidytext" # for boxplot order
)

@@

# Biogeographic regions (from Rey 2020)
Biogeo_reg <-fread(paste(my_path("D"),"Climate/SPATIAL_PARTITION_OCEAN.csv", sep ="")) %>% 
  clean_names() %>% 
  rowid_to_column("index")

@@

Eez_biogeo <- EEZ_CellID %>% 
  # filter(eezid == 191) %>% 
  left_join(Biogeo_reg,
       by = "index") %>% 
  # Deal with NaNs in data
  mutate(biome = ifelse(biome == "NaN",0,biome)) %>% 
  filter(biome > 0) %>% 
  # Get frequency of biome per index
  group_by(eezid,biome) %>% 
  summarise(nbiom=n()) %>% 
  # Select the more frequent biome per eez
  group_by(eezid) %>% 
  top_n(1,nbiom) %>% 
  ungroup() %>% 
  select(-nbiom) %>% 
  distinct(eezid, .keep_all = T) # Croatia is 1/2. We arbritary allocated the first one

@@

# Per Stock
stock_d <-  Eme_Res %>% 
  left_join(Neighbours) %>% 
  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
  group_by(taxon_key,tresh) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(tresh) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "stock"
  )

@@

#### Fig. 5 Demersal Vs Pelagic

@@

@@

reg_diff_data <- Eme_Res %>% 
  filter(tresh == "Tresh_one",
         !is.na(un_sub_region)
         )

Reg_aov <- aov(emerging_yr~un_sub_region, data = reg_diff_data)
Reg_aov

@@

hab_diff_data <- Eme_Res %>% 
  filter(tresh == "Tresh_one",
         !is.na(un_sub_region)
         ) %>% 
  left_join(Spp_env_pref,
            by = "taxon_key")

@@

### Tab.1 Explain the ToE

@@

Lm_data <- Eme_Res %>% # 19728
  filter(tresh == "Tresh_one") %>% # 11490
  left_join(Spp_depth) %>% 
  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel) %>% 
  left_join(Toe_temp,
            by = c("eezid","tresh")
            ) %>% 
  left_join(Catch_trend,
            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
            ) %>% # 11492
  left_join(Eez_biogeo,
            by = "eezid") %>% 
  left_join(Spp_env_pref,
            by = "taxon_key") %>% 
  left_join(Eez_centroids,
            by = "eezid") %>% 
  mutate(latitude = abs(latitude))

@@

# Model 1; all variables
M_all_var <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)

all_sum <- summary(M_all_var)
all_aic <- AIC(M_all_var)
all_r <-  all_sum$adj.r.squared

@@

tibble(
  "model" = paste("M",seq(1,8,1),sep=""),
  "AIC" = c(all_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic),
  "rsquared" = c(all_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r)
)

# UNIDAS 20282659

# The Winner is M8!

stargazer(M8, title = "Results", align=TRUE)

## Troubleshooting

>>>>

#### Project's Library
packages <- c(
  "dbemImport",
  "tidyverse",
  "here", # for dbem_import `here()`
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  "geosphere", # estimate distances between points `distm()`
  "ggrepel",
  "zoo", # for average mean
  "parallel", # for mclapply,
  "sf", # for mapping
  "st", # for mapping
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  "zeallot", # for Juanito's map
  "gmt", # for estimating distances between points
  "viridis",
  "cowplot",
  "rfishbase",
  "stargazer", # For nice lm tables
  "tidytext", # for boxplot order
  "moments", # for skewness and kurto
  "pgirmess" # for non parametric post hoc test
)

@@

# Biogeographic regions (from Rey 2020)
Biogeo_reg <-fread(paste(my_path("D"),"Climate/SPATIAL_PARTITION_OCEAN.csv", sep ="")) %>% 
  clean_names()

# Fix coordinate system and re-name biomes
Biogeo_reg <- suppressWarnings(Biogeo_reg[order(Biogeo_reg$lat, rev(Biogeo_reg$lon),decreasing=TRUE), ]) %>% 
  mutate(index = seq(1,259200,1)) %>% 
  mutate(
    # biome = ifelse(biome2 == 1, "c_tropical",
    #                ifelse(biome2 == 2,"c_temperate",
    #                       ifelse(biome2 == 3, "c_polar",
    #                              ifelse(biome2 == 4, "o_tropical",
    #                                     ifelse(biome2 == 5, "o_temperate",
    #                                            ifelse(biome2 == 6, "o_polar",
    #                                                   biome2
    #                                            )
    #                                     )
    #                              ) 
    #                       )
    #                )
    # ),
    biome = ifelse(biome2 %in% c(1,4), "tropical",
                   ifelse(biome2  %in% c(2,5),"temperate",
                          ifelse(biome2  %in% c(3,6), "polar",
                                 biome2)
                   )
    )
    ) %>% 
  filter(biome2 != "NaN")

# biome2 1 = Coastal tropical 
# biome2 2 = Coastal sub-tropical
# biome2 3 = Coastal polar
# biome2 4 = Oceanic tropical
# biome2 5 = Oceanic sub-tropical
# biome2 6 = Oceanic polar

# Biogeo_reg %>%
  # gather("biome_type","value",biome:biome2) %>%
# ggplot() +
# geom_tile(
#   aes(
#     x = lon,
#     y = lat,
#     fill = biome
#   )
# )
#   facet_wrap(~biome_type,
#              ncol=1)

@@

Eez_biogeo <- EEZ_CellID %>% 
  # filter(eezid == 8) %>%
  left_join(Biogeo_reg,
       by = "index") %>% 
  # Deal with NaNs in data
  mutate(biome = ifelse(biome == "NaN",0,biome)) %>% 
  filter(biome > 0) %>% 
  # Get frequency of biome per index
  group_by(eezid,biome) %>% 
  summarise(nbiom=n()) %>% 
  # Select the more frequent biome per eez
  group_by(eezid) %>% 
  top_n(1,nbiom) %>% 
  ungroup() %>% 
  select(-nbiom) %>% 
  distinct(eezid, .keep_all = T) # Croatia is 1/2. We arbritary allocated the first one

@@

# Per Stock
stock_da <-  Eme_Res %>% 
  left_join(Neighbours) %>% 
  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
  group_by(taxon_key,tresh) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(tresh) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "stock"
  )

@@

#### Fig. 5 Boxplot

@@

#### Test for normallity

```{r, normality_test, eval = F, echo = F}
# Test for normality

Stats_data <- Eme_Res %>% 
  filter(tresh == "Tresh_one",
         n_ens == 10)

# ------------------- #
# Test normal asumptions 
# ------------------- #

hist(Stats_data$emerging_yr) # Positevely skewed
qqnorm(Stats_data$emerging_yr) # not normally distributed
skewness(Stats_data$emerging_yr) #1.02
kurtosis(Stats_data$emerging_yr) #2.82

# ------------------- #

# (NOT NORMAL DISTRIRBUTION)

@@

Reg_diff_data <- Eme_Res %>% 
  filter(tresh == "Tresh_one",
         !is.na(un_sub_region)
         )

# ------------------- #
# Run non-parametric Krustal
# ------------------- #

Reg_kw <- kruskal.test(un_sub_region~emerging_yr, data = Reg_diff_data)

Reg_kw
# Kruskal-Wallis rank sum test
# 
# data:  un_sub_region by emerging_yr
# Kruskal-Wallis chi-squared = 287.23, df = 93, p-value < 2.2e-16

kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data)

# Multiple differences but Eastern Asia, Lat. Ame. Car. and Polinesia are the nes who differe the most (e.g. differe with more revions).

# ------------------- #
# Just to double check ANOVA results (give the same)
# ------------------- #
Reg_aov <- aov(emerging_yr~un_sub_region, data = Reg_diff_data)
Reg_aov

@@

Hab_diff_data <- Eme_Res %>% 
  filter(tresh == "Tresh_one",
         !is.na(un_sub_region)
         ) %>% 
  left_join(Spp_env_pref,
            by = "taxon_key")

# ------------------- #
# Run non-parametric Krustal
# ------------------- #

Hab_kw <- kruskal.test(DemersPelag~emerging_yr, data = Hab_diff_data)

Hab_kw
# Kruskal-Wallis rank sum test

# data:  DemersPelag by emerging_yr
# Kruskal-Wallis chi-squared = 286.48, df = 93, p-value < 2.2e-16

kruskalmc(emerging_yr~DemersPelag, data = Hab_diff_data)

# Multiple differences but main differences for pelagic-ocenaic and reefs

# ------------------- #
# Just to double check ANOVA results (give the same)
# ------------------- #

@@

## Table.1 Multy linear model

@@

Lm_data <- Eme_Res %>% # 19728
  filter(tresh == "Tresh_one") %>% # 11490
  left_join(Spp_depth) %>% 
  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel,un_region= un_geo_region) %>% 
  left_join(Toe_temp,
            by = c("eezid","tresh")
            ) %>% 
  left_join(Catch_trend,
            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
            ) %>% # 11492
  left_join(Eez_biogeo,
            by = "eezid") %>% 
  left_join(Spp_env_pref,
            by = "taxon_key") %>% 
  left_join(Eez_centroids,
            by = "eezid") %>% 
  mutate(latitude = abs(latitude))

@@

# Model 1; all variables
M1 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)

M1_sum <- summary(M1)
M1_aic <- AIC(M1)
M1_r <-  M1_sum$adj.r.squared

@@

# Model 9; Include UN regions
M9 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude + un_region, data = Lm_data)

M9_sum <- summary(M9)
M9_aic <- AIC(M9)
M9_r <-  M9_sum$adj.r.squared

tibble(
  "model" = paste("M",seq(1,9,1),sep=""),
  "AIC" = c(M1_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic, M9_aic),
  "rsquared" = c(M1_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r,M9_r)
)

# The Winner is M9    49283.   0.0765

#Table for paper
M9_sum %>% 
  tidy() %>% 
  filter(p.value < 0.05) %>% 
  mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks. Only showing explanatory variables. The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
            # align=TRUE)
  # knitr::kable() %>% 
  write_csv(.,
            "./Tables/mlr_result.csv")

#Table for supplements
M9_sum %>% 
  tidy() %>% 
  # filter(p.value < 0.05) %>% 
  # mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks. Only showing explanatory variables. The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
            # align=TRUE)
  # knitr::kable() %>% 
  write_csv(.,
            "./Tables/mlr_result_sup.csv")
```

# END PAPER

# Troubleshooting
